A Formal Verification Framework for Smart Contracts Based on Symbolic Execution

### 2024-01-16 papers/static-checker/ecoop2024/ecoop2024.tex
general revise

\subsection{Entrypoint relations}
\label{sec:entrypoint-relations}
For a smart contract on the blockchain, the order of smart contract
calls can be crucial. We specify ordering constraints among calls with
an entrypoint relation. An entrypoint relation can be specified as
\lstinline/%a with cd -> %b/, which starts with the name
\lstinline/%a/ of the entrypoint and is followed by the keyword
\lstinline/with/ in case conditions are specified as predicates. The
following symbol \lstinline/->/ indicates that from the entrypoint
that satisfies the condition \lstinline/cd/, it is possible  to call
the entrypoint \lstinline/%b/. One smart contract specification can
contain several entrypoint relations concatenated by
\lstinline/|/. There is a special entrypoint named \lstinline/create/
that indicates the contract state after deployment.

@@

\begin{figure}[tp]
    \centering
    \includegraphics[width=0.8\textwidth]{auction-life-cycle}
    \caption{Auction contract life cycle}
    \label{fig:auction-life-cycle}
\end{figure}

@@

The architecture of the static checker is illustrated in
Figure~\ref{fig:architecture-of-static-checker}. The system comprises
three phases. The first phase is user specification, where users
provide the Michelson code of the contract and define the properties
that the smart contract should satisfy. Subsequently, the symbolic
interpreter is run on the code with the provided input term.  The
path conditions are then fed into a Z3 converter, which transforms
symbolic terms into Z3 formulas. Concurrently, the property
specification, specifically pre- and post conditions, also undergoes
conversion to Z3 formulas. The results of the symbolic execution and
Z3 formulas are then input into the static checker, which verifies
whether these properties hold for the smart contract
implementation. This constitutes the verification phase.

@@

\subsubsection{Output check}
\label{sec:output-check}
The first check is the output check, where the checker verifies
whether the output term matches. The final state contains a stack
with only one element as \Stack\ \STACKCONCAT\ \EMPTYSTACK. There are
two different forms of the final state. In the case where the code is
given as incomplete smart contract code, the element \Stack\ contains
just a normal term and its corresponding type as (\Term,
\TY). However, in the case of a complete smart contract code, it has
the form (\PAIR\ \VOPERATIONLIST\ \VSTORAGE, \TPAIR\ (\TOPERATIONLIST)
\TYS) \STACKCONCAT\ \EMPTYSTACK, where \VSTORAGE\ is the term
representing the storage and \VOPERATIONLIST\ repesents a operation
list. Given an output term $ \tau$, the checker scans all the final
states, which do not contain a failwith stack and checks whether the
output term matches. In the former case, it matches $ \tau$  with
the term \Term. In the latter case, it essentially matches with the
corresponding part \VSTORAGE\ in the final states resulting from the
symbolic interpreter. When the input term is a concrete value, the
result from the interpreter is also concrete, and then the checker can
be used to verify if the output aligns with the expectations
specified in the output field.

Listing~\ref{lst:add-contract-specification} contains an example of an
incomplete smart contract code. The specification states that the
output of the code must be a stack that contains only one element of
type \TINT\, symbolically represented as a variable \VZ\ of type
\TINT.  In this example, there are two final states, the first one
being a failwith state recording the failwith condition when the first
element of the pair is less than the second element. The failwith
state contains the message \lstinline|'Unexpected Pair'|, which
describes the reason for the failure. The second state is a reachable
state that contains, in the top element, an integer value, which is
the sum of the two elements of the pair. The output check succeeds for
this example as expected. However, this test could fail if, for
example, the output term is incorrectly specified as a variable of
a type that does not match with the actual result. If the input
term is concrete, for instance, a pair of numbers as in
\lstinline|Pair 5 3|, then the expected output should be the number
\lstinline|8|. To check this,  the output term may be specified as
a concrete value.

@@

\subsubsection{Fail condition}
\label{sec:fail-condition}
A Michelson contract can fail due to a programmer-induced condition
caused by the instruction \FAILWITH. It terminates contract execution
with an error message reported back to the caller. This error message
includes the top value on the stack. These failures are important in
smart contract programming to guarantee that a smart contract call
should not be successful under specific circumstances. Smart contract
calls are expected to fail due to malicious behavior, such as an
unauthorized caller or invalid arguments passed along the call. For
example, the auction contract should not be closed by someone who is
not the owner of the contract, and the bidding should not succeed if
the bid is not strictly higher than the current highest bid. Several
smart contract standards, such as
\cite{@ha-supply-a-smart-contract-standard}, require the contract to fail
with specific messages under certain conditions. Therefore, it is
essential to check the fail conditions of a contract.

The static checker provides a feature to generate all fail conditions
implemented in a smart contract code. The tool searches for all
failwith states, resulting from an execution terminated by the
\FAILWITH\ instruction, from the final state space, collects the path
conditions and the recorded message. The result returned to users
includes the fail condition together with the message. For example,
for the specification in Listing \ref{lst:add-contract-specification},
the fail condition is recorded as
\lstinline/(x < y): 'Unexpected Pair'/.

@@

The auction contract exhibits several fail conditions. For the
entrypoint \lstinline/bid/, the fail condition is as shown in
Listing~\ref{lst:fail-condition-auction-contract-bid} and for the
\lstinline/close/ entrypoint, it is recorded in
Listing~\ref{lst:fail-condition-auction-contract-close}. For the
\lstinline/bid/ entrypoint, there are three fail conditions.  (1) If
the auction is closed, (2) if the auction is open, but the amount less
than or equal to the balance, (3) if the auction is open, the amount
is greater than the balance, but the highest bidder's address is
invalid. The last failure can happen because there are valid Michelson
addresses, which are not associated with any real account on the Tezos
blockchain, or the parameter's type is not correct: the transfer can
only succeed if the address is associated with an account and the
parameter supplied matches with the parameter expected by the Tezos
blockchain. Bidding returns the previous highest bid as part of the
transaction where the highest bidder is stored as a value of type
address. However, to receive a token transfer, this address must be
cast (by the Michelson implementation of the contract) to an implicit
contract of Michelson type (\TCONTRACT\ \TUNIT). This cast is
unavoidable, as a value of contract type cannot be stored.  However,
the cast may fail for the reasons just discussed, and its failure
leads to an error condition that is reported with a \FAILWITH\
instruction in the implementation. In our framework, this cast failure
shows up  as a fail condition, but the user's preconditions assume
that casting is successful. The \lstinline/close/ entrypoint has
similar fail conditions when the auction is closed or the owner's
address is cannot be cast. The other fail condition occurs when the
auction is open, but the sender who calls this entrypoint is not the
owner of the contract, which was stored in the contract owner part.
\atha{This discussion suggests that the specification is incomplete:
  it should contain the assertion that the highest bidder as well as
  the contract owner field are valid and point to a unit contract!}

\begin{lstlisting}[float=tp,captionpos=b,caption={Fail condition for auction contract bid entrypoint },label={lst:fail-condition-auction-contract-bid},numbers=left]
1. not (auction_open): 'Already closed'
2. auction_open && (Amount <= Balance - Amount):
   'Not higher than the current highest bid'
3. auction_open && not (Amount <= Balance - Amount) && 
   get_contract (unit, highest_bidder) = None:
   'Invalid highest bidder address' 
\end{lstlisting}

\begin{lstlisting}[float=tp,captionpos=b,caption={Fail condition for auction contract close entrypoint},label={lst:fail-condition-auction-contract-close},numbers=left]
1. not (auction_open): 'Already closed'
2. auction_open && Sender != contract_owner): 'Not owner'
3. auction_open && (Sender = contract_owner) &&
   get_contract (unit, contract_owner) = None:
  'Invalid contract owner address'
\end{lstlisting}

@@

The static checker enables users to verify smart contract properties
by providing pre- and post-condition specifications. A property is
specified with the clause:  \lstinline/pre-condition := &&/
$\Phi_{i}$ followed by \lstinline/post-condition := &&/ $\Psi_{j}$. This
specification asserts that if all preconditions $\Phi_{i}$ hold before
smart contract execution, then all $\Psi_{j}$ must hold after
successful (i.e., without failure) termination. To verify this
property, we explore all the paths $\Pi$ 
with the path condition $\Theta$ such that ($\bigwedge \Phi_{i}$
$\land$ $\Theta$) is satisfiable. We then check that for every path
$\Pi$, $\neg$ ($\bigwedge \Phi_{i}$ $\land$ $\Theta$ $\rightarrow \bigwedge \Psi_{j}$) is
unsatisfiable in the context given by the input for the entry point
and its corresponding final state.
\atha{Don't you have to include $\Theta$ in the check for unsat?}

Consider the specification of the sub contract in Listing
\ref{lst:sub-contract-specification}. In this context, the two
variables \lstinline/x/ and \lstinline/y/ are integers. The results of
the symbolic execution are either a failwith state with the path
condition \lstinline/x/$<$\lstinline/y/,  or a state in which the
final stack has only one element, which is a symbolic variable
\lstinline/sbvar/ of type integer, with the path condition stating
that the variable \lstinline/x/ is greater than or equal to
\lstinline/y/ and the symbolic variable \lstinline/sbvar/ is equal to
the substract of \lstinline/x/ and \lstinline/y/.  For this specification,
the failwith branch is discarded because the precondition is not
satisfy with the	 path condition, and then \lstinline/z/ is matched to
the symbolic variable \lstinline/sbvar/ in the other branch. In this
example, the checker then verifies that $\neg$ ((\lstinline/x/ >= y) $\land$ (\lstinline/z/
= \lstinline/x/ - \lstinline/y/) $\rightarrow$ (\lstinline/z/ = \lstinline/x/ - \lstinline/y/) $\land$ (\lstinline/z/ >= 0)) is
unsatisfiable.
\atha{I don't understand the last sentence/formula. Where does the
  $z\ge8$ come from?}

\subsubsection{Life-cycle check}
\label{sec:life-cycle-check}
\atha{To be completed}
An entrypoint relation can be defined as \lstinline/%a with cd -> %b/, indicating that after invoking the entrypoint \lstinline/%a/, it is permissible to call the entrypoint \lstinline/%b/ if the condition \lstinline/cd/ is satisfied. The presence of the keyword \lstinline/not/ at the beginning signals an impossibility. Note that \lstinline/with cd/ may be omitted when no condition is needed.

Let $S_a$ represent the set of all reachable states resulting from the execution of the call to the entrypoint \lstinline/%a/. Each $s_{i}$ in $S_a$ is associated with a path condition $\Theta_i$ and is represented as a stack containing a single element—a pair consisting of an operation list ($\VOPERATIONLIST_i$) and the storage $\VSTORAGE_i$. For each $s_{i}$, symbolic execution is performed on the entrypoint \lstinline/%b/ using input as a pair of the parameter and the storage $\VSTORAGE_i$, constrained by $\Theta_i$ and the condition \lstinline/cd/. In the execution state space, a state $s_{j}$ with the path condition $\Theta_j$ is considered reachable if ($\Theta_i$ $\land$ \lstinline/cd/ $\land$ $\Theta_j$) is satisfiable, and $\neg$ ($\Theta_i$ $\land$ \lstinline/cd/ $\rightarrow$ $\Theta_j$) is unsatisfiable. Let $S_b$ denote the set of these reachable states. If $S_b$ is empty or contains only failwith states, then it is impossible to invoke the entrypoint \lstinline/%b/ from the state $s_{i}$. Otherwise, it is considered a success. Let $R$ be the set containing all such states $s_{i}$ in $S_a$ where it is successful to call the entrypoint \lstinline/%b/ with the condition \lstinline/cd/. If $R$ is non-empty, then \lstinline/%a with cd -> %b/ holds. Otherwise, it is false, signifying \lstinline/not (%a with cd -> %b)/.

@@

In a blockchain ecosystem, tokens represent digital assets that can be
transferred between accounts. These tokens are securely stored and
verified within the blockchain network. There are various types of
tokens, including native tokens (cryptocurrencies) like Bitcoin,
Ether, Tezos, and others. Additionally, tokens can be created as
overlays through smart contracts on an existing blockchain. Two common
categories are: (1) Fungible Tokens (FT), which are interchangeable
and (2) Non-Fungible Tokens (NFTs), which represent unique digital
assets, and each unit is distinct. Stablecoins, falling under fungible
tokens, are designed to maintain a stable value by tracking the value
of a fiat currency like USD or EUR.
\atha{add citations!}

While designers can create tokens that behave in any way they want, it
is advisable to implement tokens that adhere to established standards.
This adherence guarantees interoperability and compatibility across
different platforms. A token standard defines a set of rules and an
interface that smart contracts must follow to ensure compatibility
with decentralized exchanges, wallets, and other tools within the
ecosystem. For example, ERC-20 is a token standard used in Ethereum.

\atha{provide references for the standards!}
The Tezos blockchain has its own set of standards known as Financial
Applications (FA). Two notable standards are: (1) FA 1.2, designed for
fungible tokens, and (2) FA 2, a more versatile standard accommodating
both fungible and non-fungible tokens (NFTs). The standard outlines a
smart contract that functions as a ledger, associating identities with
balances. This ledger is designed to facilitate token transfer
operations and enable approvals for spending tokens from other
accounts. Any contract aiming to implement the standard must
incorporate the functions specified in
Listing~\ref{lst:fa12-interface} such as token transfer (transactions
between users), approval (allowing one user to send tokens from
another user's address),  allowance (browsing the amount of tokens
available for an account to sending from another address), user
balance (checking the balance of a user), and current supply (checking
the total supply of tokens across all addresses), where
\lstinline/view a r = pair a (contract r)/ implies an view entrypoint.
A view is an entrypoint that takes an
argument of type \lstinline/a/ and a callback that results in a value
of type \lstinline/r/. The callback in this case will be another
contract address. Further elaboration on the details of FA 1.2 can be
found in \cite{financial-applications-1.2}, and we will delve into
these functions in subsequent 
sections.
\atha{is that correct: ``callback that results in a value of type r''?
I would have expected that the callback \textbf{expects} a value of
type r.}

@@

In this section, we verify USDtz \cite{USDtz-implementation}, an
implementation of the FA 1.2 standard. Our goal is to ensure that the
implementation aligns with the standard, verifying its
correctness. USDtz is a stablecoin and one of the first and
longest-running coins on the Tezos blockchain, launched in September
2020. The smart contract is located at the address
\texttt{KT1LN4LPSqTMS7Sd2CJw4bbDGRkMv2t68Fy9}, and the code can be
found on the Tezos blockchain \cite{USDtz-address}. USDtz adheres to
the FA token standard, and the initial deployment of the USDtz smart
contract follows the FA 1.2 standard. 
The implementation consists of approximately 1000 instructions. It
contains all five entrypoints required in the FA 1.2 interface and
four additional ones.  We will verify each entrypoint of this smart
contract. First, we examine how the approvable ledger is implemented
in the USDtz smart contract. The ledger is a substantial map that
links user addresses to a pair consisting of their balance and a map
associating addresses with the amount of tokens allowed to be
transferred from that user's
account. Listing~\ref{lst:storage-udstz-contract} shows the storage type of the
USDtz smart contract.

@@

\subsubsection{transfer}
As illustrated in Listing \ref{lst:fa12-interface}, the
\lstinline/transfer/ entrypoint credits the account associated with
the address provided in the \lstinline/to/ parameter, while debiting
the account corresponding to the \lstinline/from/ parameter with the
specified \lstinline/value/ amount. The FA 1.2 standard mandates that
this call must fail under the following two conditions: (1) In the
event of insufficient balance, this entrypoint MUST fail with the
error \lstinline/NotEnoughBalance/, indicating insufficient funds in
the sender's account to execute the transfer. (2) If the sender's
account has no permission to withdraw the \lstinline/value/ amount of
funds, the call will fail with the \lstinline/NotEnoughAllowance/
error, which also contains the same pair message as the first failure
scenario. To verify these requirements, we utilize the failwith
condition feature of the tool and examine the returned results, as
outlined in Listing~\ref{lst:transfer-udstz-contract}.
\begin{lstlisting}[float,captionpos=b,caption={Fail conditions for the \lstinline/transfer/ entrypoint},label={lst:transfer-udstz-contract},numbers=left]
1. paused: 'TokenOperationsArePaused'
2. not paused &&
   get_big_map (from, ledger) = None : 'NotEnoughBalance'
3. not paused &&
   get_big_map (from, ledger) = Some sbvar &&
   sbvar = Pair sbvar1 sbvar2 &&
   sbvar1 < value : 'NotEnoughBalance'
4. not paused &&
   get_big_map (from, ledger) = Some sbvar &&
   sbvar = Pair sbvar1 sbvar2 &&
   get_map (Sender, sbvar2) = Some sbvar3 &&
   sbvar3 < value: 'NotEnoughAllowance'
\end{lstlisting}
The fundamental property that this entrypoint should satisfy is the decrement of the allowance of the sender's account by the \lstinline/value/ that has been sent as Listing \ref{lst:specification-transfer}.
\begin{lstlisting}[float,captionpos=b,caption={Specification of the \lstinline/transfer/ entrypoint},label={lst:specification-transfer},numbers=left]
mcontract usdtz = spec 
   storage :=  Pair (ledger: big_map address 
      (pair nat (map address nat))) 
      (Pair (admin: address) 
            (Pair (paused: bool) (total_supply: nat)))
   entrypoint %transfer
      parameter := Pair (from: address)
         (Pair (to: address) (value: nat))
      pre-condition := not paused 
        && get_big_map (from, ledger) = 
           Some (v: option (pair nat (map address nat))) 
        && (v = Pair (n1: nat) (m: map address nat))
        && get_map (Sender, m) = Some (n2: nat) 
        && (n1 >= value) && (n2 >= value)
      post-condition := get_big_map (from, post_ledger) = 
           Some (post_v: option (pair nat (map address nat))) 
        && post_v = Pair (post_n1: nat) (post_m: map address nat) 
        && get_map (Sender, post_m) = Some (post_n2: nat) 
        && (post_n1 = n1 - value) && (post_n2 = n2 - value)
\end{lstlisting}

>>>>

\subsection{Entrypoint relations}
\label{sec:entrypoint-relations}
For a smart contract on the blockchain, the order of smart contract
calls can be crucial. We specify ordering constraints among calls with
an entrypoint relation. An entrypoint relation can be specified as
\lstinline/%a -> %b with cd/, which starts with the name
\lstinline/%a/ of the entrypoint and is followed by the keyword
\lstinline/with/ in case conditions are specified as predicates. The
following symbol \lstinline/->/ indicates that from the entrypoint
that satisfies the condition \lstinline/cd/, it is possible  to call
the entrypoint \lstinline/%b/. One smart contract specification can
contain several entrypoint relations concatenated by
\lstinline/|/. There is a special entrypoint named \lstinline/create/
that indicates the contract state after deployment.

@@

\begin{figure}[tp]
    \centering
    \includegraphics[width=1.1\textwidth]{auction-life-cycle}
    \caption{Auction contract life cycle}
    \label{fig:auction-life-cycle}
\end{figure}

@@

%The architecture of the static checker is illustrated in Figure~\ref{fig:architecture-of-static-checker}. 
The system comprises three phases. The first phase is user specification, where users
provide the Michelson code of the contract and define the properties
that the smart contract should satisfy. Subsequently, the symbolic
interpreter is run on the code with the provided input term.  The
path conditions are then fed into a Z3 converter, which transforms
symbolic terms into Z3 formulas. Concurrently, the property
specification, specifically pre- and post conditions, also undergoes
conversion to Z3 formulas. The results of the symbolic execution and
Z3 formulas are then input into the static checker, which verifies
whether these properties hold for the smart contract
implementation. This constitutes the verification phase.

@@

\subsubsection{Output check}
\label{sec:output-check}
The first check is the output check, where the checker verifies
whether the output term matches. The final state contains a stack
with only one element as \Stack\ \STACKCONCAT\ \EMPTYSTACK. There are
two different forms of the final state. In the case where the code is
given as incomplete smart contract code, the element \Stack\ contains
just a normal term and its corresponding type as (\Term,
\TY). However, in the case of a complete smart contract code, it has
the form (\PAIR\ \VOPERATIONLIST\ \VSTORAGE, \TPAIR\ (\TOPERATIONLIST)
\TYS) \STACKCONCAT\ \EMPTYSTACK, where \VSTORAGE\ is the term
representing the storage and \VOPERATIONLIST\ repesents a operation
list. Given an output term $ \tau$, the checker scans all the reachable final
states, which do not contain a failwith stack and checks whether the
output term matches. In the former case, it matches $ \tau$  with
the term \Term. In the latter case, it essentially matches with the
corresponding part \VSTORAGE\ in the final states resulting from the
symbolic interpreter. When the input term is a concrete value, the
result from the interpreter is also concrete, and then the checker can
be used to verify if the output aligns with the expectations
specified in the output field.

Listing~\ref{lst:sub-contract-specification} contains an example of an
incomplete smart contract code. The specification states that the
output of the code must be a stack that contains only one element of
type \TINT\, symbolically represented as a variable \VZ\ of type
\TINT.  In this example, there are two final states, the first one
being a failwith state recording the failwith condition when the first
element of the pair is less than the second element. The failwith
state contains the message \lstinline|'Unexpected Pair'|, which
describes the reason for the failure. The second state is a reachable
state that contains, in the top element, an integer value, which is
the substract of the two elements of the pair. The output check succeeds for
this example as expected. However, this test could fail if, for
example, the output term is incorrectly specified as a variable of
a type that does not match with the actual result. If the input
term is concrete, for instance, a pair of numbers as in
\lstinline|Pair 5 3|, then the expected output should be the number
\lstinline|2|. To check this,  the output term may be specified as
a concrete value.

@@

\subsubsection{Fail condition}
\label{sec:fail-condition}
A Michelson contract can fail due to a programmer-induced condition
caused by the instruction \FAILWITH. It terminates contract execution
with an error message reported back to the caller. This error message
includes the top value on the stack. These failures are important in
smart contract programming to guarantee that a smart contract call
should not be successful under specific circumstances. Smart contract
calls are expected to fail due to malicious behavior, such as an
unauthorized caller or invalid arguments passed along the call. For
example, the auction contract should not be closed by someone who is
not the owner of the contract, and the bidding should not succeed if
the bid is not strictly higher than the current highest bid. Several
smart contract standards, such as
\cite{erc, fa}, require the contract to fail
with specific messages under certain conditions. Therefore, it is
essential to check the fail conditions of a contract.

The static checker provides a feature to generate all fail conditions
implemented in a smart contract code. The tool searches for all
failwith states, resulting from an execution terminated by the
\FAILWITH\ instruction, from the final state space, collects the path
conditions and the recorded message. The result returned to users
includes the fail condition together with the message. For example,
for the specification in Listing \ref{lst:sub-contract-specification},
the fail condition is recorded as
\lstinline/(x < y): 'Unexpected Pair'/.

@@

The auction contract exhibits several fail conditions. For the
entrypoint \lstinline/bid/, the fail condition is as shown in
Listing~\ref{lst:fail-condition-auction-contract-bid} and for the
\lstinline/close/ entrypoint, it is recorded in
Listing~\ref{lst:fail-condition-auction-contract-close}. For the
\lstinline/bid/ entrypoint, there are three fail conditions.  (1) If
the auction is closed, (2) if the auction is open, but the amount less
than or equal to the balance, (3) if the auction is open, the amount
is greater than the balance, but the highest bidder's address is
invalid. The last failure can happen because there are valid Michelson
addresses, which are not associated with any real account on the Tezos
blockchain, or the parameter's type is not correct: the transfer can
only succeed if the address is associated with an account and the
parameter supplied matches with the parameter expected by the Tezos
blockchain. Bidding returns the previous highest bid as part of the
transaction where the highest bidder is stored as a value of type
address. However, to receive a token transfer, this address must be
cast (by the Michelson implementation of the contract) to an implicit
contract of Michelson type (\TCONTRACT\ \TUNIT). This cast is
unavoidable, as a value of contract type cannot be stored.  However,
the cast may fail for the reasons just discussed, and its failure
leads to an error condition that is reported with a \FAILWITH\
instruction in the implementation. The \lstinline/close/ entrypoint has
similar fail conditions when the auction is closed or the owner's
address is cannot be cast. The other fail condition occurs when the
auction is open, but the sender who calls this entrypoint is not the
owner of the contract, which was stored in the contract owner part.

\begin{lstlisting}[float=tp,captionpos=b,caption={Fail condition for auction contract bid entrypoint },label={lst:fail-condition-auction-contract-bid},numbers=left]
1. not (auction_open = true): 'Already closed'
2. (auction_open = true) && (Amount <= pre_Balance):
   'Not higher than the current highest bid'
3. (auction_open = true) && (Amount > pre_Balance) && 
   get_contract (unit, highest_bidder) = None:
   'Invalid highest bidder address' 
\end{lstlisting}

\begin{lstlisting}[float=tp,captionpos=b,caption={Fail condition for auction contract close entrypoint},label={lst:fail-condition-auction-contract-close},numbers=left]
1. not (auction_open = true): 'Already closed'
2. (auction_open = true) && (Sender != contract_owner): 
    'Not the owner'
3. (auction_open = true) && (Sender = contract_owner) &&
   get_contract (unit, contract_owner) = None:
  'Invalid contract owner address'
\end{lstlisting}

@@

The static checker enables users to verify smart contract properties
by providing pre- and post-condition specifications. A property is
specified with the clause:  \lstinline/pre-condition := &&/
$\Phi_{i}$ followed by \lstinline/post-condition := &&/ $\Psi_{j}$. This
specification asserts that if all preconditions $\Phi_{i}$ hold before
smart contract execution, then all $\Psi_{j}$ must hold after
successful (i.e., without failure) termination. To verify this
property, we explore all the paths $\Pi$ 
with the path condition $\Theta$ such that ($\bigwedge \Phi_{i}$
$\land$ $\Theta$) is satisfiable. We then check that for every path
$\Pi$, $\neg$ ($\bigwedge \Phi_{i}$ $\land$ $\Theta$ $\rightarrow \bigwedge \Psi_{j}$) is unsatisfiable in the context given by the input for the entry point
and its corresponding final state.

Consider the specification of the sub contract in Listing
\ref{lst:sub-contract-specification}. In this context, the two
variables \lstinline/x/ and \lstinline/y/ are integers. The results of
the symbolic execution are either a failwith state with the path
condition \lstinline/x/$<$\lstinline/y/,  or a state in which the
final stack has only one element, which is a symbolic variable
\lstinline/sbvar/ of type integer, with the path condition stating
that the variable \lstinline/x/ is greater than or equal to
\lstinline/y/ and the symbolic variable \lstinline/sbvar/ is equal to
the substract of \lstinline/x/ and \lstinline/y/.  For this specification,
the failwith branch is discarded because the precondition is not
satisfy with the	 path condition, and then \lstinline/z/ is matched to
the symbolic variable \lstinline/sbvar/ in the other branch. In this
example, the checker then verifies that $\neg$ ((\lstinline/x/ >= y) $\land$ (\lstinline/z/
= \lstinline/x/ - \lstinline/y/) $\rightarrow$ (\lstinline/z/ = \lstinline/x/ - \lstinline/y/) $\land$ (\lstinline/z/ >= 0)) is unsatisfiable.

\subsubsection{Life-cycle check}
\label{sec:life-cycle-check}
An entrypoint relation can be defined as \lstinline/(%a -> %b) with cd/, indicating that after invoking the entrypoint \lstinline/%a/, it is permissible to call the entrypoint \lstinline/%b/ if the condition \lstinline/cd/ is satisfied. The presence of the keyword \lstinline/not/ at the beginning signals an impossibility. Note that \lstinline/with cd/ may be omitted when no condition is needed.

Let $S_a$ represent the set of all reachable states resulting from the execution of the call to the entrypoint \lstinline/%a/. Each $s_{i}$ in $S_a$ is associated with a path condition $\Theta_i$ and is represented as a stack containing a single element—a pair consisting of an operation list ($\VOPERATIONLIST_i$) and the storage $\VSTORAGE_i$. For each $s_{i}$, symbolic execution is performed on the entrypoint \lstinline/%b/ using input as a pair of the parameter and the storage $\VSTORAGE_i$, constrained by $\Theta_i$ and the condition \lstinline/cd/. In the execution state space, a state $s_{j}$ with the path condition $\Theta_j$ is considered reachable if ($\Theta_i$ $\land$ \lstinline/cd/ $\land$ $\Theta_j$) is satisfiable, and $\neg$ ($\Theta_i$ $\land$ \lstinline/cd/ $\rightarrow$ $\Theta_j$) is unsatisfiable. Let $S_b$ denote the set of these reachable states. If $S_b$ is empty or contains only failwith states, then it is impossible to invoke the entrypoint \lstinline/%b/ from the state $s_{i}$. Otherwise, it is considered a success. Let $R$ be the set containing all such states $s_{i}$ in $S_a$ where it is successful to call the entrypoint \lstinline/%b/ with the condition \lstinline/cd/. If $R$ is non-empty, then \lstinline/(%a -> %b) with cd/ holds. Otherwise, it is false, signifying \lstinline/not (%a  -> %b) with cd/.

@@

In a blockchain ecosystem, tokens represent digital assets that can be
transferred between accounts. These tokens are securely stored and
verified within the blockchain network. There are various types of
tokens, including native tokens (cryptocurrencies) like Bitcoin,
Ether, Tezos, and others. Additionally, tokens can be created as
overlays through smart contracts on an existing blockchain. Two common
categories are: (1) Fungible Tokens (FT), which are interchangeable
and (2) Non-Fungible Tokens (NFTs), which represent unique digital
assets, and each unit is distinct. Stablecoins, falling under fungible
tokens, are designed to maintain a stable value by tracking the value
of a fiat currency like USD or EUR.

While designers can create tokens that behave in any way they want, it
is advisable to implement tokens that adhere to established standards.
This adherence guarantees interoperability and compatibility across
different platforms. A token standard defines a set of rules and an
interface that smart contracts must follow to ensure compatibility
with decentralized exchanges, wallets, and other tools within the
ecosystem. For example, ERC-20 \cite{erc} is a token standard used in Ethereum.

The Tezos blockchain has its own set of standards known as Financial
Applications (FA). Two notable standards are: (1) FA 1.2 \cite{fatzip}, designed for
fungible tokens, and (2) FA 2 \cite{fatzips}, a more versatile standard accommodating
both fungible and non-fungible tokens (NFTs). The standard outlines a
smart contract that functions as a ledger, associating identities with
balances. This ledger is designed to facilitate token transfer
operations and enable approvals for spending tokens from other
accounts. Any contract aiming to implement the standard must
incorporate the functions specified in
Listing~\ref{lst:fa12-interface} such as token transfer (transactions
between users), approval (allowing one user to send tokens from
another user's address),  allowance (browsing the amount of tokens
available for an account to sending from another address), user
balance (checking the balance of a user), and current supply (checking
the total supply of tokens across all addresses), where
\lstinline/view a r = pair a (contract r)/ implies an view entrypoint.
A view is an entrypoint that takes an
argument of type \lstinline/a/ and a callback that results in a value
of type \lstinline/r/. The callback in this case will be another
contract address. Further elaboration on the details of FA 1.2 can be
found in \cite{usdtz}, and we will delve into
these functions in subsequent 
sections.

@@

In this section, we verify USDtz \cite{usdtzimplemetation}, an
implementation of the FA 1.2 standard. Our goal is to ensure that the
implementation aligns with the standard, verifying its
correctness. USDtz is a stablecoin and one of the first and
longest-running coins on the Tezos blockchain, launched in September
2020. The smart contract is located at the address
\texttt{KT1LN4LPSqTMS7Sd2CJw4bbDGRkMv2t68Fy9}, and the code can be
found on the Tezos blockchain \cite{tzstatsusdtz}. USDtz adheres to
the FA token standard, and the initial deployment of the USDtz smart
contract follows the FA 1.2 standard. 
The implementation consists of approximately 1000 instructions. It
contains all five entrypoints required in the FA 1.2 interface and
four additional ones.  We will verify each entrypoint of this smart
contract. First, we examine how the approvable ledger is implemented
in the USDtz smart contract. The ledger is a substantial map that
links user addresses to a pair consisting of their balance and a map
associating addresses with the amount of tokens allowed to be
transferred from that user's
account. Listing~\ref{lst:storage-udstz-contract} shows the storage type of the
USDtz smart contract.

@@

\subsubsection{transfer}
As illustrated in Listing \ref{lst:fa12-interface}, the
\lstinline/transfer/ entrypoint credits the account associated with
the address provided in the \lstinline/to/ parameter, while debiting
the account corresponding to the \lstinline/from/ parameter with the
specified \lstinline/value/ amount. The FA 1.2 standard mandates that
this call must fail under the following two conditions: (1) In the
event of insufficient balance, this entrypoint MUST fail with the
error \lstinline/NotEnoughBalance/, indicating insufficient funds in
the sender's account to execute the transfer. (2) If the sender's
account has no permission to withdraw the \lstinline/value/ amount of
funds, the call will fail with the \lstinline/NotEnoughAllowance/
error, which also contains the same pair message as the first failure
scenario. To verify these requirements, we utilize the failwith
condition feature of the tool and examine the returned results, as
outlined in Listing~\ref{lst:transfer-udstz-contract}\footnote{There may be some duplicates of the same failwith condition due to the branching of the symbolic interpreter. The results here are shortened. We are working on filtering them in the tool.}.
\begin{lstlisting}[float,captionpos=b,caption={Fail conditions for the \lstinline/transfer/ entrypoint},label={lst:transfer-udstz-contract},numbers=left]
1. (paused = true): 'TokenOperationsArePaused'
2. not (paused = true) &&
   get_big_map (from, ledger) = None : 'NotEnoughBalance'
3. not (paused = true) &&
   get_big_map (from, ledger) = Some sbvar &&
   sbvar = Pair sbvar1 sbvar2 &&
   sbvar1 < value : 'NotEnoughBalance'
4. not (paused = true) &&
   get_big_map (from, ledger) = Some sbvar &&
   sbvar = Pair sbvar1 sbvar2 &&
   get_map (Sender, sbvar2) = Some sbvar3 &&
   sbvar3 < value: 'NotEnoughAllowance'
\end{lstlisting}
The fundamental property that this entrypoint should satisfy is the decrement of the allowance of the sender's account by the \lstinline/value/ that has been sent as Listing \ref{lst:specification-transfer}.
\begin{lstlisting}[float,captionpos=b,caption={Specification of the \lstinline/transfer/ entrypoint},label={lst:specification-transfer},numbers=left]
mcontract usdtz = spec 
   storage :=  Pair (ledger: big_map address 
      (pair nat (map address nat))) 
      (Pair (admin: address) 
            (Pair (paused: bool) (total_supply: nat)))
   entrypoint %transfer
      parameter := Pair (from: address)
         (Pair (to: address) (value: nat))
      pre-condition := not paused 
        && get_big_map (from, ledger) = 
           Some (v: pair nat (map address nat))
        && (v = Pair (n: nat) (m: map address nat))
        && get_map (Sender, m) = Some (k: nat) 
        && (n >= value) && (k >= value)
      post-condition := get_big_map (from, post_ledger) = 
           Some (post_v: option (pair nat (map address nat))) 
        && post_v = Pair (post_n: nat) (post_m: map address nat) 
        && get_map (Sender, post_m) = Some (post_k: nat) 
        && (post_n = n - value) && (post_k = k - value)
\end{lstlisting}
